Expert-In-The-Loop Causal Discovery: Iterative Model Refinement Using Expert Knowledge

### 2025-02-08 paper/main.tex
Updates screenshot
Updates

\begin{figure}[t!]
	\centering
	\begin{subfigure}{0.5\textwidth}
		\centering
		\includegraphics{../code/fig3/shd_ribbon.pdf}
		\caption{}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
		\centering
		\includegraphics{../code/fig3/sid_ribbon.pdf}
		\caption{}
	\end{subfigure}
	\caption{Comparison of PC and Hill-Climb Search algorithms against
		manually drawn DAGs using the assistance method. As PC and
		Hill-Climb Search return the Markov Equivalence Class (MEC), we
		use the best and worst scoring orientation of the MEC to get
		the range of values. The human drawn values are done for
		multiple accuracy values of the oracle.}
\end{figure}

\begin{figure}[t!]
	\begin{subfigure}{0.25\textwidth}
		\centering
		\includegraphics{../code/fig4/unexplained_effect.pdf}
		\caption{}
	\end{subfigure}%
	\begin{subfigure}{0.25\textwidth}
		\centering
		\includegraphics{../code/fig4/ll.pdf}
		\caption{}
	\end{subfigure}
	\caption{Plots showing improvements in fit of the model over iterations
	on the Adult income dataset. For the expert simulation an LLM is used
	for the edge orientations. (a) Shows using total unexplained effect that is the
	sum of effect size between variables that don't have an edge between them. (b)
	Shows the log-likelihood of the overall model.
	}
\end{figure}

@@

@@

Another major difference is in the interpretability of the evaluation metric.
Most scoring metrics are based on log-likelihood with a penalty for model
complexity. These score metrics allow for relative comparisons between models
but do not provide an absolute measure of model fit. That is, they indicate
which model is better for a given dataset but do not reveal how well the model
explains the data in an absolute sense. In contrast, summing our measure of
association across all variables provides a direct fitness measure—this sum
approaches zero when the model perfectly accounts for all observed correlations
in the data. This property enables an absolute assessment of model quality,
rather than just a relative comparison between models.

@@

\begin{figure}[t!]
	\centering
	\includegraphics[scale=0.4]{../code/plots/web_tool_full.png}
	\caption{A screenshot of the web tool for constructing the model. Users
		can upload their dataset after which the tool creates an empty
		graph and shows all pair of variables which are associated in
		the model using undirected red edges with the strength of
		association represented using edge width. Users can then
		iteratively add edges to the model (shown in green) while
		deciding the edge orientation based on domain knowledge.
		Unnecessary edges are shown in black.}
	\label{fig:web}
\end{figure}

>>>>

\begin{figure}[t!]
	\centering
	\begin{subfigure}{0.5\textwidth}
		\centering
		\includegraphics{../code/fig3/shd_ribbon.pdf}
		\caption{}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
		\centering
		\includegraphics{../code/fig3/sid_ribbon.pdf}
		\caption{}
	\end{subfigure}
	\caption{Comparison of PC, Hill-Climb Search, and GES algorithms against
		\textit{ExpertInLoop} algorithm. As automated algorithms only recover
		the CPDAG, we use the best and worst scoring orientation of the CPDAG 
		to get the range. The \textit{ExpertInLoop} shows values for various
		$ \alpha_{\textit{eff}} $ values corresponding to $ \alpha = \{0.1, 0.3, 0.5, 0.7, 0.9\} $.}
	\label{fig:shd_sid}
\end{figure}

\begin{figure}[t!]
	\begin{subfigure}{0.25\textwidth}
		\centering
		\includegraphics{../code/fig4/unexplained_effect.pdf}
		\caption{}
	\end{subfigure}%
	\begin{subfigure}{0.25\textwidth}
		\centering
		\includegraphics{../code/fig4/ll.pdf}
		\caption{}
	\end{subfigure}
	\caption{Plots showing improvements in fit of the model over iterations
	on the Adult income dataset. For the expert simulation an LLM is used
	for the edge orientations. (a) Shows using total unexplained effect that is the
	sum of effect size between variables that don't have an edge between them. (b)
	Shows the log-likelihood of the overall model.
	}
	\label{fig:unexplained_ll}
\end{figure}

@@

\begin{equation}
	\alpha_{\mathrm{eff}} = \alpha + (1 - \alpha) / 3
\end{equation}

@@

Another major difference is in the interpretability of the evaluation metric.
Most scoring metrics are based on log-likelihood with a penalty for model
complexity. These score metrics allow for relative comparisons between models
but do not provide an absolute measure of model fit. That is, they indicate
which model is better for a given dataset but do not reveal how well the model
explains the data in an absolute sense. In contrast, summing our measure of
association across all variables provides a direct fitness measure—this sum
approaches zero when the model perfectly accounts for all observed correlations
in the data. This property enables an absolute assessment of model quality,
rather than just a relative comparison between models. Figure~\ref{fig:unexplained_ll}

@@

\begin{figure}[t!]
	\centering
	\includegraphics[scale=0.4]{../code/plots/web_tool_full_new.png}
	\caption{A screenshot of the web tool for constructing the model. Users
		can upload their dataset after which the tool creates an empty
		graph and shows all pair of variables which are associated in
		the model using undirected red edges with the strength of
		association represented using edge width. Users can then
		iteratively add edges to the model (shown in green) while
		deciding the edge orientation based on domain knowledge.
		Unnecessary edges are shown in black.}
	\label{fig:web}
\end{figure}
